---
title: "Data Manipulation"
author: "Yaxuan Deng"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

Let's import the datasets

```{r}
litters_df <- 
  read_csv("data_import_examples/FAS_litters.csv",na=c("NA",".",""))
litters_df <- 
  janitor::clean_names(litters_df)

pups_df <- 
  read_csv("data_import_examples/FAS_pups.csv",skip = 3,na = c("NA",".",""))
pups_df <- 
  janitor::clean_names(pups_df)
```
## select

'select' selects variables from my dataframes

```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```

select a range

```{r}
select(litters_df,group:gd_of_birth)
```

select by removal

```{r}
select(litters_df,-group)
select(litters_df,-group:gd_of_birth)
```

select by prefix

```{r}
select(litters_df,group,starts_with("gd"))
```

rearrange by selecting

```{r}
select(litters_df,starts_with("gd"),group)
select(litters_df,litter_number,everything())
```

```{r}
relocate(litters_df,litter_number)
```


rename by selecting

```{r}
select(litters_df,GROUP = group,everything())
rename(litters_df,GROUP = group)
```
what if i only select one variable?

```{r}
select(litters_df,group)
pull(litters_df,group)
```

```{r}
select(pups_df, sex,pd_ears)
```

## filter

use 'filter()' to filter out rows

```{r}
filter(litters_df,gd_of_birth == 20, pups_born_alive == 6)
```

look for more than two pups alive

```{r}
filter(litters_df,pups_born_alive > 5)
filter(litters_df,pups_born_alive >= 5)
filter(litters_df,pups_born_alive <= 5)
```

look for non-matches

```{r}
filter(litters_df,pups_born_alive != 6)
```

match to a set

```{r}
filter(litters_df,group %in% c("Con7","Con8"))
```

filter out missing rows with 'drop_na()'

```{r}
drop_na(litters_df)
drop_na(litters_df,gd0_weight)
```
```{r}
filter(pups_df,sex == 1)
```


## mutate

use mutate() to create or modify variables

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## arrange

it arranges things

```{r}
arrange(litters_df,pups_born_alive)
arrange(litters_df,group,pups_born_alive)
```

## pipes

a bad way to so things

```{r}
litters_df <- 
  read_csv("data_import_examples/FAS_litters.csv",na=c("NA",".",""))
litters_df <- 
  janitor::clean_names(litters_df)

litters_df2 <- select(litters_df,group,litter_number,starts_with("gd"))

litters_df3 <- drop_na(litters_df)

litters_df4 <- mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight
)

# 4 is actually what I'm interested in
```

instead, use pip to avoid all these
command+shift+m

```{r}
litters_df <- 
  read_csv("data_import_examples/FAS_litters.csv",na=c("NA",".","")) %>% 
  janitor::clean_names() %>% 
  select(,group,litter_number,starts_with("gd")) %>% 
  drop_na() %>% 
  mutate(
    wt_gain = gd18_weight - gd0_weight
  )
```

